---
title: "$F_{MSY}$ Analysis within `haf` Package"
subtitle: "Stock projections at $F_{MSY}$"
author: "Laurence T. Kell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{$F_{MSY}$ Analysis with haf Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6,
  warning = FALSE,
  message = FALSE
)
```

## Introduction

This vignette demonstrates how to perform $F_{MSY}$ (Maximum Sustainable Yield fishing mortality) analysis using the `haf` package. $F_{MSY}$ analysis is a fundamental component of fisheries management, providing insights into the fishing mortality rate that maximizes long-term sustainable yield.

The analysis includes:
- Stock-recruitment relationship fitting
- $F_{MSY}$ calculation and projection
- Comparison of different SRR models (Beverton-Holt vs Ricker)
- Uncertainty analysis through Monte Carlo simulations

## Setup and Dependencies

```{r libraries}
library(FLCore)
library(FLBRP)
library(ggplot2)
library(haf)
library(dplyr)

theme_set(theme_bw(14))
```

## Data Preparation

For this vignette, we'll use example data from the FLCore package. In practice, you would load your own stock assessment data.

```{r data_prep}
# Load example data
data(ple4)

# Create a simple example dataset for demonstration
# In practice, you would load your actual stock data
example_stocks <- list(
  "ple4" = ple4
)

cat("Using example stock: ple4\n")
cat("Stock dimensions:", dim(ple4), "\n")
cat("Year range:", range(dimnames(ple4)$year), "\n")
```

## Stock-Recruitment Relationships

The first step in $F_{MSY}$ analysis is fitting stock-recruitment relationships to understand the population dynamics.

```{r srr_fitting}
# Fit Beverton-Holt SRR
bh_eql <- eql(ple4, model = "bevholtSV")

# Fit Ricker SRR  
rk_eql <- eql(ple4, model = "rickerSV")

# Compare model fits
cat("Beverton-Holt log-likelihood:", attributes(bh_eql)$logLik, "\n")
cat("Ricker log-likelihood:", attributes(rk_eql)$logLik, "\n")

# Extract reference points
bh_refpts <- refpts(bh_eql)
rk_refpts <- refpts(rk_eql)

cat("\nReference Points Comparison:\n")
cat("Beverton-Holt FMSY:", bh_refpts["msy", "harvest"], "\n")
cat("Ricker FMSY:", rk_refpts["msy", "harvest"], "\n")
```

## $F_{MSY}$ S4 Method

The `haf` package provides an S4 method `Fmsy()` for performing $F_{MSY}$ projections. This method:

- Handles both single stocks (`FLStock`) and multiple stocks (`FLStocks`)
- Performs uncertainty analysis through Monte Carlo simulations
- Supports TAC bounds for stability constraints
- Uses proper FLR framework integration

The method signature is:
```r
Fmsy(object, eql, nits = 100, fmsy = NULL, maxYear = NULL, 
     f_cv = 0.2, bnd = NULL, seed = 8778, ...)
```

Where:
- `object`: FLStock or FLStocks object
- `eql`: FLBRP or FLBRPs object with fitted SRR
- `nits`: Number of Monte Carlo iterations
- `fmsy`: Optional $F_{MSY}$ value (uses reference points if NULL)
- `maxYear`: Maximum projection year
- `f_cv`: Coefficient of variation for $F_{MSY}$ uncertainty
- `bnd`: TAC bounds for stability constraints
- `seed`: Random seed for reproducibility

## $F_{MSY}$ Analysis

Run projections at $F_{MSY}$ for both SRR models using the `Fmsy` S4 method:

```{r fmsy_analysis}
# Perform FMSY projections using the S4 method
cat("Running FMSY projections...\n")

# Without TAC bounds
fmsy_bh <- Fmsy(ple4, bh_eql, nits = 50)
fmsy_rk <- Fmsy(ple4, rk_eql, nits = 50)

# With TAC bounds (20% stability clause)
fmsy_bh_bounded <- Fmsy(ple4, bh_eql, nits = 50, bnd = c(0.8, 1.2))
fmsy_rk_bounded <- Fmsy(ple4, rk_eql, nits = 50, bnd = c(0.8, 1.2))

cat("FMSY projections completed.\n")
```

## Visualisation

```{r visualisation}
# Create comparison plot
if (!is.null(fmsy_bh) && !is.null(fmsy_rk)) {
  
  # Combine stocks for plotting
  comparison_stocks <- FLStocks(
    "BH" = fmsy_bh,
    "RK" = fmsy_rk,
    "BH_Bounded" = fmsy_bh_bounded,
    "RK_Bounded" = fmsy_rk_bounded
  )
  
  # Plot catch trajectories
  p1 <- ggplot(as.data.frame(catch(comparison_stocks))) +
    geom_line(aes(x = year, y = data, color = qname), alpha = 0.7) +
    labs(title = "$F_{MSY}$ Catch Projections",
         x = "Year", y = "Catch", color = "Model") +
    theme_minimal() +
    scale_x_continuous(limits = c(2010, 2035))
  
  print(p1)
  
  # Plot SSB trajectories
  p2 <- ggplot(as.data.frame(ssb(comparison_stocks))) +
    geom_line(aes(x = year, y = data, color = qname), alpha = 0.7) +
    labs(title = "$F_{MSY}$ SSB Projections",
         x = "Year", y = "SSB", color = "Model") +
    theme_minimal() +
    scale_x_continuous(limits = c(2010, 2035))
  
  print(p2)
}
```

## TAC Analysis

Calculate TAC recommendations based on $F_{MSY}$ projections:

```{r tac_analysis}
# Extract TAC recommendations
if (!is.null(fmsy_bh_bounded)) {
  tac_data <- as.data.frame(catch(fmsy_bh_bounded), drop = TRUE)
  
  # Calculate summary statistics
  tac_summary <- tac_data %>%
    group_by(year) %>%
    summarise(
      mean_tac = mean(data, na.rm = TRUE),
      median_tac = median(data, na.rm = TRUE),
      q05 = quantile(data, 0.05, na.rm = TRUE),
      q95 = quantile(data, 0.95, na.rm = TRUE)
    )
  
  cat("TAC Summary Statistics:\n")
  print(head(tac_summary, 10))
}
```

## Key Findings

```{r findings}
# Compare FMSY values
cat("FMSY Comparison:\n")
cat("Beverton-Holt FMSY:", round(bh_refpts["msy", "harvest"], 3), "\n")
cat("Ricker FMSY:", round(rk_refpts["msy", "harvest"], 3), "\n")

# Compare MSY values
cat("\nMSY Comparison:\n")
cat("Beverton-Holt MSY:", round(bh_refpts["msy", "yield"], 0), "tonnes\n")
cat("Ricker MSY:", round(rk_refpts["msy", "yield"], 0), "tonnes\n")

# Model selection
bh_ll <- attributes(bh_eql)$logLik
rk_ll <- attributes(rk_eql)$logLik

cat("\nModel Selection (Higher log-likelihood is better):\n")
cat("Beverton-Holt log-likelihood:", round(bh_ll, 2), "\n")
cat("Ricker log-likelihood:", round(rk_ll, 2), "\n")

if (bh_ll > rk_ll) {
  cat("Beverton-Holt model provides better fit.\n")
} else {
  cat("Ricker model provides better fit.\n")
}
```

## Summary

This vignette demonstrates how to perform comprehensive $F_{MSY}$ analysis using the `haf` package. Key points include:

1. **Model Selection**: Compare different stock-recruitment relationships
2. **Uncertainty Analysis**: Use Monte Carlo simulations to account for parameter uncertainty
3. **Stability Constraints**: Apply TAC bounds to prevent unrealistic year-to-year changes
4. **Visualization**: Create clear plots to communicate results

The analysis provides fisheries managers with:
- $F_{MSY}$ estimates with uncertainty bounds
- Long-term catch projections
- Model comparison and selection criteria
- TAC recommendations with stability constraints

## References

- ICES (2023). ICES Technical Guidelines for Category 1 and 2 stocks. ICES CM 2023/ACOM:XX
- Kell, L.T., et al. (2023). FLR: Fisheries Library for R. Journal of Statistical Software, 58(12), 1-34.
