---
title: "Management Strategy Evaluation (MSE) with the haf Package"
subtitle: "A practical guide to using the MSE method for fisheries management"
author: "haf package"
date: "`r format(Sys.time(),'%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    highlight: tango
mathjax: TRUE
tags: [FLR, FLCore, MSE, fisheries management, harvest control rules]
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  cache = TRUE,
  fig.width = 10,
  fig.height = 6,
  dev = "png",
  comment = NA
)

# Load required packages
library(FLCore)
library(ggplotFL)
library(FLBRP)
library(FLasher)
library(FLAssess)
library(haf)
library(plyr)
library(dplyr)
library(ggplot2)
library(patchwork)
library(doParallel)
library(foreach)

# Set theme for plots
theme_set(theme_bw(14))
```

# Introduction

Management Strategy Evaluation (MSE) is a framework for testing the performance of harvest control rules (HCRs) under various sources of uncertainty. The `haf` package provides a generic `MSE()` method that implements a "shortcut" MSE approach, which uses historical recruitment residuals and observation errors to simulate future scenarios.

This vignette demonstrates how to use the `MSE()` method for both single stocks and multiple stocks, showing how to:

1. Set up stock-recruitment relationships
2. Configure MSE parameters
3. Run MSE simulations
4. Analyze and visualize results

# Single Stock MSE Example

## Data Preparation

First, let's load some example data and prepare it for MSE analysis:

```{r data_prep}
# Load example stock data
data(ple4)

# Display basic stock information
summary(ple4)

# Plot the stock
plot(ple4) + 
  ggtitle("Plaice Stock Data") +
  theme(plot.title = element_text(hjust = 0.5))
```

## Fitting Stock-Recruitment Relationships

The MSE method requires fitted stock-recruitment relationships. Let's fit both Beverton-Holt and Ricker models:

```{r srr_fitting}
# Fit Beverton-Holt stock-recruitment relationship
eql_bh <- eql(ple4, model = "bevholtSV")

# Fit Ricker stock-recruitment relationship  
eql_rk <- eql(ple4, model = "rickerSV")

# Compare model fits
logLik_bh <- attributes(eql_bh)$logLik[1]
logLik_rk <- attributes(eql_rk)$logLik[1]

cat("Beverton-Holt log-likelihood:", logLik_bh, "\n")
cat("Ricker log-likelihood:", logLik_rk, "\n")
cat("Best model:", ifelse(logLik_bh > logLik_rk, "Beverton-Holt", "Ricker"), "\n")

# Plot stock-recruitment relationships
sr_plot <- ggplot() +
  geom_point(data = as.data.frame(attributes(eql_bh)$rec.obs), 
             aes(x = ssb, y = rec), alpha = 0.6) +
  geom_line(data = as.data.frame(predict(attributes(eql_bh)$sr)), 
            aes(x = ssb, y = rec, color = "Beverton-Holt"), size = 1) +
  geom_line(data = as.data.frame(predict(attributes(eql_rk)$sr)), 
            aes(x = ssb, y = rec, color = "Ricker"), size = 1) +
  labs(x = "Spawning Stock Biomass", y = "Recruitment", 
       title = "Stock-Recruitment Relationships",
       color = "Model") +
  theme(plot.title = element_text(hjust = 0.5))

print(sr_plot)
```

## Running MSE for Single Stock

Now let's run MSE using the Beverton-Holt model:

```{r single_stock_mse}
# Run MSE with default parameters
mse_result <- MSE(ple4, eql_bh, 
                  nits = 50,        # Reduced for vignette
                  endYear = 2030,   # Shorter projection period
                  seed = 123)

# Check if MSE was successful
if (!is.null(mse_result)) {
  cat("MSE completed successfully!\n")
  cat("Number of iterations:", dim(mse_result)[6], "\n")
  cat("Projection years:", range(dimnames(mse_result)$year), "\n")
} else {
  cat("MSE failed - check if FMSY is available in benchmark\n")
}
```

## Analyzing MSE Results

Let's analyze the MSE results and create some visualizations:

```{r analyze_results}
if (!is.null(mse_result)) {
  # Extract key metrics
  ssb_mse <- ssb(mse_result)
  catch_mse <- catch(mse_result)
  fbar_mse <- fbar(mse_result)
  
  # Calculate summary statistics
  ssb_summary <- apply(ssb_mse, c(1, 2), quantile, probs = c(0.05, 0.25, 0.5, 0.75, 0.95))
  catch_summary <- apply(catch_mse, c(1, 2), quantile, probs = c(0.05, 0.25, 0.5, 0.75, 0.95))
  
  # Plot results
  p1 <- plot(ssb_mse) + 
    ggtitle("Spawning Stock Biomass") +
    theme(plot.title = element_text(hjust = 0.5))
  
  p2 <- plot(catch_mse) + 
    ggtitle("Catch") +
    theme(plot.title = element_text(hjust = 0.5))
  
  p3 <- plot(fbar_mse) + 
    ggtitle("Fishing Mortality") +
    theme(plot.title = element_text(hjust = 0.5))
  
  # Combine plots
  combined_plot <- p1 / p2 / p3
  print(combined_plot)
  
  # Calculate performance metrics
  final_ssb <- ssb_mse[, ac(2030)]
  final_catch <- catch_mse[, ac(2030)]
  
  cat("Final year (2030) summary:\n")
  cat("SSB median:", median(final_ssb), "thousand tonnes\n")
  cat("SSB 90% CI:", quantile(final_ssb, c(0.05, 0.95)), "thousand tonnes\n")
  cat("Catch median:", median(final_catch), "thousand tonnes\n")
  cat("Catch 90% CI:", quantile(final_catch, c(0.05, 0.95)), "thousand tonnes\n")
}
```

# Multiple Stocks MSE Example

## Creating Multiple Stocks

Let's create a simple example with multiple stocks by using the same stock with different parameters:

```{r multiple_stocks}
# Create multiple stocks with slight modifications
ple4_high <- ple4
ple4_low <- ple4

# Modify natural mortality for different scenarios
m(ple4_high) <- m(ple4_high) * 1.2  # Higher natural mortality
m(ple4_low) <- m(ple4_low) * 0.8    # Lower natural mortality

# Create FLStocks object
stocks <- FLStocks(
  "Base" = ple4,
  "High_M" = ple4_high,
  "Low_M" = ple4_low
)

# Create corresponding equilibrium objects
eqls <- FLBRPs(
  "Base" = eql_bh,
  "High_M" = eql(ple4_high, model = "bevholtSV"),
  "Low_M" = eql(ple4_low, model = "bevholtSV")
)

# Display stock information
cat("Multiple stocks created:\n")
for (i in names(stocks)) {
  cat(i, ": SSB range =", range(ssb(stocks[[i]])), "thousand tonnes\n")
}
```

## Running MSE for Multiple Stocks

Now let's run MSE for all stocks simultaneously:

```{r multiple_stocks_mse}
# Run MSE for multiple stocks
mse_multi <- MSE(stocks, eqls, 
                 nits = 30,        # Reduced for vignette
                 endYear = 2030,   # Shorter projection period
                 seed = 456)

# Check results
if (length(mse_multi) > 0) {
  cat("MSE completed for", length(mse_multi), "stocks\n")
  
  # Compare final year results
  final_results <- data.frame(
    Stock = names(mse_multi),
    SSB_median = sapply(mse_multi, function(x) median(ssb(x)[, ac(2030)])),
    SSB_cv = sapply(mse_multi, function(x) cv(ssb(x)[, ac(2030)])),
    Catch_median = sapply(mse_multi, function(x) median(catch(x)[, ac(2030)])),
    Catch_cv = sapply(mse_multi, function(x) cv(catch(x)[, ac(2030)]))
  )
  
  print(final_results)
} else {
  cat("MSE failed for all stocks\n")
}
```

## Visualizing Multiple Stock Results

Let's create comparative visualizations:

```{r multi_visualization}
if (length(mse_multi) > 0) {
  # Create FLStocks from MSE results for easy plotting
  mse_stocks <- FLStocks(mse_multi)
  
  # Plot SSB comparison
  ssb_plot <- plot(ssb(mse_stocks)) + 
    ggtitle("Spawning Stock Biomass Comparison") +
    theme(plot.title = element_text(hjust = 0.5))
  
  # Plot catch comparison
  catch_plot <- plot(catch(mse_stocks)) + 
    ggtitle("Catch Comparison") +
    theme(plot.title = element_text(hjust = 0.5))
  
  # Combine plots
  print(ssb_plot / catch_plot)
  
  # Create performance comparison table
  performance_table <- data.frame(
    Stock = names(mse_multi),
    SSB_2030_median = round(sapply(mse_multi, function(x) median(ssb(x)[, ac(2030)])), 1),
    SSB_2030_cv = round(sapply(mse_multi, function(x) cv(ssb(x)[, ac(2030)])), 3),
    Catch_2030_median = round(sapply(mse_multi, function(x) median(catch(x)[, ac(2030)])), 1),
    Catch_2030_cv = round(sapply(mse_multi, function(x) cv(catch(x)[, ac(2030)])), 3)
  )
  
  cat("Performance comparison (final year 2030):\n")
  print(performance_table)
}
```

# Advanced MSE Configuration

## Custom HCR Parameters

You can specify custom harvest control rule parameters:

```{r custom_hcr}
# Define custom HCR parameters
custom_hcr <- FLPar(
  ftar = 0.3,      # Target fishing mortality
  btrig = 200000,  # Biomass trigger (tonnes)
  fmin = 0.01,     # Minimum fishing mortality
  bmin = 50000,    # Minimum biomass
  blim = 25000     # Limit biomass
)

# Run MSE with custom parameters
mse_custom <- MSE(ple4, eql_bh, 
                  hcrParams = custom_hcr,
                  nits = 20,
                  endYear = 2030,
                  seed = 789)

if (!is.null(mse_custom)) {
  cat("Custom HCR MSE completed successfully!\n")
  
  # Compare with default HCR
  if (!is.null(mse_result)) {
    default_catch <- median(catch(mse_result)[, ac(2030)])
    custom_catch <- median(catch(mse_custom)[, ac(2030)])
    
    cat("Default HCR median catch (2030):", round(default_catch, 1), "tonnes\n")
    cat("Custom HCR median catch (2030):", round(custom_catch, 1), "tonnes\n")
    cat("Difference:", round(custom_catch - default_catch, 1), "tonnes\n")
  }
}
```

## Parallel Processing Configuration

The MSE method uses parallel processing by default. You can control the number of cores:

```{r parallel_config}
# Run MSE with specific number of cores
mse_parallel <- MSE(ple4, eql_bh, 
                    cores = 2,      # Use only 2 cores
                    nits = 20,
                    endYear = 2030,
                    seed = 999)

cat("MSE with 2 cores completed:", !is.null(mse_parallel), "\n")
```

# Performance Analysis

## Risk Analysis

Let's analyze the risk of stock depletion:

```{r risk_analysis}
if (!is.null(mse_result)) {
  # Calculate risk metrics
  ssb_mse <- ssb(mse_result)
  
  # Risk of SSB falling below 20% of current level
  current_ssb <- median(ssb(ple4)[, ac(dims(ple4)$maxyear)])
  threshold <- current_ssb * 0.2
  
  risk_by_year <- apply(ssb_mse, 2, function(x) mean(x < threshold))
  
  # Plot risk over time
  risk_data <- data.frame(
    Year = as.numeric(names(risk_by_year)),
    Risk = risk_by_year
  )
  
  risk_plot <- ggplot(risk_data, aes(x = Year, y = Risk)) +
    geom_line(size = 1, color = "red") +
    geom_hline(yintercept = 0.05, linetype = "dashed", color = "orange") +
    labs(title = "Risk of SSB < 20% of Current Level",
         x = "Year", y = "Probability") +
    theme(plot.title = element_text(hjust = 0.5))
  
  print(risk_plot)
  
  cat("Risk analysis summary:\n")
  cat("Final year risk:", round(risk_by_year[length(risk_by_year)], 3), "\n")
  cat("Years with risk > 5%:", sum(risk_by_year > 0.05), "\n")
}
```

## Stability Analysis

Let's analyze the stability of catch recommendations:

```{r stability_analysis}
if (!is.null(mse_result)) {
  catch_mse <- catch(mse_result)
  
  # Calculate catch stability (coefficient of variation)
  catch_cv <- apply(catch_mse, 2, cv)
  
  # Plot catch stability
  stability_data <- data.frame(
    Year = as.numeric(names(catch_cv)),
    CV = catch_cv
  )
  
  stability_plot <- ggplot(stability_data, aes(x = Year, y = CV)) +
    geom_line(size = 1, color = "blue") +
    geom_hline(yintercept = 0.2, linetype = "dashed", color = "orange") +
    labs(title = "Catch Stability (Coefficient of Variation)",
         x = "Year", y = "CV") +
    theme(plot.title = element_text(hjust = 0.5))
  
  print(stability_plot)
  
  cat("Stability analysis summary:\n")
  cat("Average catch CV:", round(mean(catch_cv), 3), "\n")
  cat("Years with CV > 20%:", sum(catch_cv > 0.2), "\n")
}
```

# Summary

This vignette has demonstrated how to use the `MSE()` method from the `haf` package for Management Strategy Evaluation. Key features include:

## Key Features

1. **Generic Design**: Works with both single stocks (`FLStock`) and multiple stocks (`FLStocks`)
2. **Flexible Parameters**: Customizable HCR parameters, projection periods, and iteration counts
3. **Parallel Processing**: Efficient computation using multiple CPU cores
4. **Error Handling**: Robust error handling with informative warnings
5. **Reproducibility**: Seed-based random number generation for reproducible results

## Typical Workflow

1. **Data Preparation**: Ensure stock data is properly formatted
2. **SRR Fitting**: Fit stock-recruitment relationships using `eql()`
3. **MSE Execution**: Run MSE with appropriate parameters
4. **Results Analysis**: Analyze performance metrics and risks
5. **Visualization**: Create plots and tables for communication

## Best Practices

- Use sufficient iterations (100+ recommended for production)
- Test multiple stock-recruitment models
- Analyze both central tendency and uncertainty
- Consider risk metrics alongside performance metrics
- Validate results against known benchmarks

The `MSE()` method provides a powerful and flexible tool for fisheries management strategy evaluation, supporting evidence-based decision making in fisheries management.

# Session Information

```{r session_info}
# Package versions
cat("Package versions:\n")
cat("FLCore:", as.character(packageVersion("FLCore")), "\n")
cat("FLBRP:", as.character(packageVersion("FLBRP")), "\n")
cat("FLasher:", as.character(packageVersion("FLasher")), "\n")
cat("haf:", as.character(packageVersion("haf")), "\n")

# R version
cat("\nR version:", R.version.string, "\n")
cat("Compiled:", date(), "\n")
```

# References

1. Kell, L.T., et al. (2007). FLR: an open-source framework for the evaluation and development of management strategies. *ICES Journal of Marine Science*, 64(4), 640-646.

2. Punt, A.E., et al. (2016). Management strategy evaluation: best practices. *Fish and Fisheries*, 17(2), 303-334.

3. ICES (2021). ICES Technical Guidelines for Management Strategy Evaluations. *ICES Cooperative Research Report*, 358.
